---
title: "EDA"
output: html_notebook
---

### The Base Plotting System

Add 1 by 1 the stuff...
title, axis labels, colors...

```{r}
library(datasets)
data(cars)
with(cars, plot(speed, dist))

```

#### Simple hist
```{r}
hist(airquality$Ozone)
```

#### Simple scatterplot
```{r}
with(airquality, plot(Wind, Ozone))
```

#### Simple boxplot
```{r}
airquality <- transform(airquality, Month = factor(Month))
boxplot(Ozone ~ Month, airquality, xlab = "Month", ylab = "Ozone (ppb)" , pch = "*")
```

#### Important Base Graphics Parameters
  pch     -the plotting symbol
  lty     -(lineetype)
  lwd     -line width
  col     -color
  xlab, ylab
  las     -oritenation of the axis labels on plot
  bg      -backgroung
  mar     -margin
  oma     -outer margin
  mfrow   -number of plots per row, col (row wise)
  mfcol   -number of plots per row, col (col wise)
  par("bg")
  plot    -make a scatter plot
  lines   - add lines to a plot
  points    - add points
  text    - add text (x,y)
  title   - add annotation
  mtext   - arbitrary text to the margin
  axis    - adding axis/ticks labels
  

#### The big One !

```{r, figures-side, fig.show="hold", out.width="50%"}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in NYC", type = "n"))
with (subset(airquality, Month == 5), points (Wind, Ozone, col = "blue"))
with (subset(airquality, Month != 5), points (Wind, Ozone, col = "red"))
legend("topright", pch = 1, col = c("blue", "red"), legend = c("May", "Other Months"))
```

```{r, figures-side, fig.show="hold", out.width="50%"}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in NYC", pch = 20))
model <- lm(Ozone ~ Wind, airquality)
abline(model, lwd = 2, col="red")
```

```{r, figures-side, fig.show="hold", out.width="10%"}
par(mfrow = c(1,3), mar = c(2,2,2,1), oma = c(0,0,2,0))
with (airquality, {
  plot(Wind, Ozone, main = "Ozone and Wind")
  abline(model, lwd = 2, col="red")
  plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
  abline(model, lwd = 2, col="red")
  plot(Temp, Ozone, main = "Ozone and Temp")
  abline(model, lwd = 2, col="red")
  mtext("Ozone and Weather in NYC", outer = T)
})
```

#### Base Plotting Demonstration
```{r, figures-side, fig.show="hold", out.width="50%"}
x <- rnorm(100)
# hist(x)
y <- rnorm(100)
# plot(x,y)
z <- rnorm (100)
# plot(x,z)
# plot(x,y)

par(mar = c(4,4,2,2))

plot(x,z, pch=4)
title("ScatterPlots")
legend("topleft", legend = "Data", pch = 20)
fit <- lm(x ~ z)
abline(fit, lw=3, col="blue")
z <- rpois(100,2)

par("mar")
#example("points")

```
```{r, figures-side, fig.show="hold", out.width="50%"}
par(mfrow = c(2,2))
plot(x, y , pch = 20)
plot(x, y , pch = 19)
plot(x, z , pch = 20)
plot(z, y , pch = 19)
```

```{r, figures-side, fig.show="hold", out.width="50%"}
x <- rnorm(100)
y <- x + rnorm(100)
g <- gl(2,50, labels = c("Male", "Female"))
maleLM <- lm(x[g == "Male"] ~ y[g == "Male"])
femaleLM <- lm(x[g == "Female"] ~ y[g == "Female"])
plot(x,y, type = "n")
points(x[g == "Male"], y[g == "Male"], col = "green")
points(x[g == "Female"], y[g == "Female"], col = "blue")
abline(maleLM, col= "green", lw=3)
abline(femaleLM, col= "blue", lw=3)
```
#### Plotting to a Device

SVG       or    bitmap
Vectors   or    Pixels
PDF             jpg


PDF
```{r, figures-side, fig.show="hold", out.width="50%"}

pdf(file = "myplot.pdf")

x <- rnorm(100)
y <- x + rnorm(100)
g <- gl(2,50, labels = c("Male", "Female"))
maleLM <- lm(x[g == "Male"] ~ y[g == "Male"])
femaleLM <- lm(x[g == "Female"] ~ y[g == "Female"])
plot(x,y, type = "n")
points(x[g == "Male"], y[g == "Male"], col = "green")
points(x[g == "Female"], y[g == "Female"], col = "blue")
abline(maleLM, col= "green", lw=3)
abline(femaleLM, col= "blue", lw=3)

dev.off()
```
```{r, figures-side, fig.show="hold", out.width="50%"}



x <- rnorm(100)
y <- x + rnorm(100)
g <- gl(2,50, labels = c("Male", "Female"))
maleLM <- lm(x[g == "Male"] ~ y[g == "Male"])
femaleLM <- lm(x[g == "Female"] ~ y[g == "Female"])
plot(x,y, type = "n")
points(x[g == "Male"], y[g == "Male"], col = "green")
points(x[g == "Female"], y[g == "Female"], col = "blue")
abline(maleLM, col= "green", lw=3)
abline(femaleLM, col= "blue", lw=3)
dev.copy(png, file = "myplot.png")
dev.off() 

```

  
## The Lattice Plotting System

xyplot, bwplot
```{r}

library(lattice)
state <- data.frame(state.x77, region = state.region)
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4,1) )

```

### The GGPLOT2 System
```{r}
library(ggplot2)
data(mpg)
qplot(displ , hwy, data = mpg)


```

