---
title: "EDA"
output: html_notebook
---



### The GGPLOT2 System
```{r}
library(ggplot2)
data(mpg)
qplot(displ , hwy, data = mpg)


```

Grammar of Graphics (leland)
Think, verbe, noun, adjective for graphics
Therory of graphics enablers

Stat graphics is a mapping of aestethic and geomteric objects

Works well with "Factors"

facets, stats, scales

## QPLOT (quick plot)

```{r, figures-side, fig.show="hold", out.width="50%"}
str(mpg)


```

## Hello World
```{r, figures-side, fig.show="hold", out.width="50%"}
qplot(displ, hwy, data=mpg)


```
  ## Modify Aesthetic
  
```{r}
qplot(displ, hwy, data = mpg, color = drv)
```
```{r}
qplot(displ, hwy, data = mpg, geom=c("point", "smooth"))
```

1 variable = Histogram
```{r}
qplot(hwy, data = mpg, fill = drv)
```
```{r}
qplot(displ, hwy, data = mpg, facets =. ~drv)
```
```{r}
qplot(hwy, data = mpg, facets =drv ~., binwidth = 1)
```
```{r}

qplot(hwy, data = mpg, fill = drv, binwidth = 1)
```
```{r}
qplot(displ, hwy, data = mpg, color = class, geom=c("point", "smooth"), method = 'lm')


``` 
```{r}
head(mpg)
```
```{r}
library(ggplot2)
g <- ggplot(mpg, aes(hwy, cty) )
summary(g)

```
```{r}
g <- ggplot(mpg, aes(cty, hwy) )
p <- g  + geom_point(aes(color = cyl))+ geom_smooth(method = "lm", size=4, linetype = 3)+ facet_grid(. ~ model)
l <- p + labs(title = "my graph", subtitle = "my graph 1 ")
print(l)

```
```{r}
library(datasets)
data(airquality)

qplot(Wind, Ozone, data = airquality, facets = . ~ factor(Month))


```
  
```{r}
airquality = transform(airquality, Month = factor(Month))
qplot(Wind, Ozone, data = airquality, facets = . ~ Month)
```
```{r}
qplot(Wind, Ozone, data = airquality, geom = "smooth")
```
```{r}
library(ggplot2)

library(ggplot2movies)
g <- ggplot(movies, aes(votes, rating))
p <- g  + geom_point(aes(color = votes)) + geom_smooth(method = "lm", size=4, linetype = 3)+ facet_grid(. ~ votes)
class(p)
print(g)

qplot(votes, rating, data = movies, smooth = "loess")
```


```{r}

g <- ggplot(mpg, aes(cty, hwy) )
p <- g  + geom_point(aes(color = cyl))+ geom_smooth(method = "lm", size=4, linetype = 3)
l <- p + labs(title = "my graph", subtitle = "my graph 1 ")
print (p)
```
```{r}

library(ggthemes) # Charger

g <- ggplot(mpg, aes(cty, hwy) )
p <- g  + geom_point(aes(color = cyl), color = "red")+ geom_smooth(method = "lm") + facet_grid(. ~ cyl)+ theme_economist()


l <- p + labs(title = "my graph", subtitle = "my graph 1 ") + xlab("Dose (mg)") + ylab("Teeth length")
print (l)
```

#### Axis Limits
```{r}
testdata <- data.frame(x = 1:100, y = rnorm(100))
testdata[50,2] <- 100 ### outlier introduced
g <- ggplot(testdata, aes(x = x, y = y))
p <- g+geom_line()
print(p)
```
#### Exclude Outlier
```{r}

p <- g + geom_line() + ylim (-3,3)
print(p)
```
#### Inculude Outlier but frame on mydata
```{r}
p <- g + geom_line() + coord_cartesian(ylim = c(-3,3))
print (p)
```
cutpoints <- quantile( mydata$xxx, seq(0,1, lenght = 4), na.rm = TRUE)

```{r}

g <- ggplot(mpg, aes(hwy, cty))
p <- g + geom_point(alpha = 1/3) + facet_wrap(.~cyl, nrow = 2, ncol = 3) + geom_smooth(method = "lm", se=FALSE, col="steelblue") + theme_bw(base_family = "Avenir", base_size = 10) + labs(x = "yoyoyo") + labs(y = "Nocturnal Symptoms") + labs(title = "MAACS Cohort")
print (p)

```
  
  