---
title: "EDA"
output: html_notebook
---

# Generic rules
## Definition

### Agglomerative approach:
- find closest two things
- put them ogtether
- find next closer

### Requires
- defined distance
- merging approach

### Produces
- tree showing how close things are to each other

### How do we deine close:
- garbage in -> garbage out
- Distance or similarity
  - euclidian
  - continuous
  - binary
  
### Example :
 - Manhattan distance


## Practice
### Generate dataset example

```{r}
set.seed(1234)
par(mar = c(0,0,0,0))
x <- rnorm(12, mean = rep(1:3, each = 4), sd=0.2)
y <- rnorm(12, mean = rep(c(1,2,1), each = 4), sd=0.2)
plot(x,y,col="blue", pch=19, cex=2)
text(x + 0.05, y + 0.05, labels=as.character(1:12))
```

### Generate dataframe
```{r}
dataFrame <- data.frame(x,y)
dist(dataFrame)
```

#### hClust
```{r}
distxy <- dist(dataFrame)
hclust <- hclust(distxy)
plot(hclust)
```

#### Prettier dendrogram
```{r}
myhclust <- function(hclust, lab = hclust$labels, lab.col=rep(1, length(hclust$labels)),
                     hang = 0.1,...){
  y <- rep(hclust$height, 2)
  x <- as.numeric(hclust$merge)
  y <- y[which(x<0)]
  x <- x[which(x<0)]
  x <- abs(x)
  y <- y[order(x)]
  x <- x[order(x)]
  plot(hclust, labels = FALSE, hang = hang, ...)
  text(x = x, y = y[hclust$order] - (max(hclust$height)*hang), labels=lab[hclust$order], col=lab.col[hclust$order], srt=90,adj=c(1,0.5), xpd = NA, ...)
}

myhclust(hclust=hclust, lab = rep(1:3, each = 4), lab.col = rep(1:3, each = 4))
```

### HeatMap

```{r}
dataMatrix <- as.matrix(dataFrame)[sample(1:12),]
heatmap(dataMatrix)

```

### K-means

Same for clusering general definiton, how do we define close...

Partitionning apporach
- Fix a number of clusters
- Get centroïds of each cluster
- Assign things to closest centroïd
- recalculate centroïds

Requires
- A defined distance metric
- numùber of clusters
- an intial gusse of centroïds location

Produces:
- Final centroid locaotion
- assignement of each points to cluster
```{r}
kmeansDF <- kmeans(dataFrame, centers = 3)
kmeansDF
```

### HeatMap
```{r}
kmeansDF <- kmeans(dataMatrix, centers = 3)
par(mfrow = c(1,2), mar=c(2,4,0.1,0.1))
image(t(dataMatrix)[, nrow(dataMatrix):1], yaxt = "n")
image(t(dataMatrix)[, order(kmeansDF$cluster)], yaxt = "n")

```

## Dimension Reduction


  
```{r}
set.seed(12345)
par(mar=rep(0.2,4))
dataMatrix1 <- matrix(rnorm(400), nrow=40)
image(1:10, 1:40, t(dataMatrix1)[,nrow(dataMatrix1):1])

```

```{r}
par(mar = rep(0.2,4))
heatmap(dataMatrix1)
```
### Add a pattern
```{r}
for (i in 1:40){
  # flip a coin
  coinFlip <- rbinom(1, size = 1, prob = 0.5)
  # if coin is head ass a common pattern to that row
  dataMatrix1[i,] <- dataMatrix1[i,] + rep(c(0,3), each = 5)
}

par(mar = rep(0.2,4))
heatmap(dataMatrix1)

```

```{r}
hh <- hclust(dist(dataMatrix1))
dataMatrix1Ordered <- dataMatrix1[hh$order,]
par(mar = rep(0.2,4))
image(t(dataMatrix1Ordered)[, nrow(dataMatrix1Ordered):1])
plot(rowMeans(dataMatrix1Ordered), 40:1,  xlab = "Row Mean", ylab = "Row", pch = 19)
plot(colMeans(dataMatrix1Ordered),xlab = "Column",  ylab = "Column mean",  pch = 19)
```

### Related problems
Get rid of unwanted/useless data which is not

### PCA/SVD
SVD Matrix decomposition